|

### 2021-07-31 prova_knit_finale.Rmd
SV modified

\begin{align}
y_{t}|x_{t} & \sim N(0,e^{\frac{x_{t}}{2}}) \\
x_{t}|x_{t-1} & \sim N(\alpha+\beta x_{t-1},\tau^2)
\end{align}

@@

In the Bootstrap Particle filter we basically use the model transition kernel and 
do resampling any time the effective sample size is smaller than a predetermined threshold.
\begin{algorithm} BPF for Stochastic Volatility Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarize $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ form $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item Draw $x_{t}^{(i)} \sim N(\alpha+\beta x_{t-1},\tau^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t-1})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{1:t-1})$
\item Set $w_{t}^{(i)} = w_{t-1}^{(i)}f_{N}(y_{t};0,e^{\frac{x_{t}}{2}}) \ \ i=1,...,N$ such that $\{(x_{t},w_{t})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{1:t})$
\item Normalize the weights: $w_{t}^{(i)}=\frac{\tilde{w}_{t}^{(i)}}{\sum_{j=1}^{N}(\tilde{w}_{t}^{(j)})}$
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $p(x_{t}|y_{1:t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{enumerate}
\end{itemize}
\end{algorithm}

If we instead want to sample from the optimal transition kernel, things get more complicated. 
It can be shown that the optimal transition kernel is
\[ m_{t}^{opt}(x_{t}|x_{t-1}) \propto f_{t}(y_{t}|x_{t})p(x_{t}|x_{t-1}) \propto exp\bigg[-\frac{1}{2\sigma^{2}}(x_{t}-\alpha-\beta x_{t-1})^{2}-\frac{e^{-x_{t}}}{2}y_{t}^{2}-\frac{x_{t}}{2}\bigg]\]
and it is not straightforward to sample from it.

\begin{algorithm} GPF for Stochastic Volatility Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarize $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ from $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item Draw $x_{t}^{(i)} \sim N(\alpha+\beta x_{t-1}+\frac{\tau^2}{4}\big[y_{t}^2e^{-(\alpha+\beta x_{t-1})}-2\big],\tau^{2}) \ \ i=1,...,N$ such that $\{(x_{t},w_{t-1})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{1:t-1})$
\item Set $w_{t}^{(i)} = w_{t-1}^{(i)}\frac{f_{N}(y_{t};0,e^{\frac{x_{t}}{2}})f_{N}(x_{t};\alpha+\beta x_{t-1},\tau^{2})}{f_{N}(x_{t};\alpha+\beta x_{t-1}+\frac{\tau^2}{4}\big[y_{t}^2e^{-(\alpha+\beta x_{t-1})}-2\big],\tau^{2})} \ \ i=1,...,N$ such that $\{(x_{t},w_{t})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{1:t})$
\item Normalize the weights: $w_{t}^{(i)}=\frac{\tilde{w}_{t}^{(i)}}{\sum_{j=1}^{N}(\tilde{w}_{t}^{(j)})}$
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $p(x_{t}|y_{1:t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{enumerate}
\end{itemize}
\end{algorithm}

\begin{algorithm}APF with optimal transition kernel for Stochastic Volatility Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ from $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw $I_{k}$ with $P(I_{k}=i) \propto w_{t-1}^{(i)}f_{N}(y_{t}|g(x_{t-1}^{(i)}),\sigma^{2}+\tau^{2}) $
\item Compute $\mu_{opt}=\mu(x_{t-1}^{(I_{k})})$
\item Draw $x_{t}^{(k)} \sim N(\mu_{opt},\sigma^{2}_{opt})$
\item Set  $\tilde{w}^{(i)}_{t} = N^{-1}$ for $i=1,...,N$
\end{enumerate}
\item otherwise
\begin{enumerate}
\item Compute $\mu_{opt}=\mu(x_{t-1})$
\item Draw $x_{t}^{(i)} \sim N(\mu_{opt},\sigma^{2}_{opt})$ for $i=1,...N$
\item Set $\tilde{w}^{(i)}_{t}=w^{(i)}_{t-1}f_{N}(y_{t};x_{t}^{(i)},\sigma^{2}+\tau^{2})$ for $i=1,...,N$
\end{enumerate}
\item Normalize the weights: $w_{t}^{(i)}=\frac{\tilde{w}_{t}^{(i)}}{\sum_{j=1}^{N}(\tilde{w}_{t}^{(j)})}$
\end{enumerate}
\end{itemize}
\end{algorithm}

\begin{algorithm} LWF for Random Walk plus Noise Model
\begin{itemize}
\item Initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ from $N(m_{0},C_{0})$, $(\alpha^{(1)},...,\alpha^{(N)})$ from $N(e_{\alpha},v_\alpha)$ and
$(\beta^{(1)},...,\beta^{(N)})$ from $N(e_{\beta},v_\beta)$ $({\tau^{2}}^{(1)},...,{\tau^{2}}^{(N)})$ from $IG(\nu_{0}/2,\nu_{0}\lambda/2)$. Set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$ and chose the appropriate value for $\delta$. Therefore $\psi^{(i)}=({\alpha}^{(i)},{\beta}^{(i)},{\tau^{2}}^{(i)})$, and
$\hat{\pi}_{0}=p(x_{0}|y_{0})=\sum_{i=1}^{N}w_{0}^{(i)}\delta_{(x_{0}^{(i)},\psi^{(i)})}$
\item For $t=1,...,n$:.
\begin{enumerate}
\item Compute $\overline{\psi}=E_{\hat{\pi}_{t-1}}(\psi)$ and $\Sigma=V_{\hat{\pi}_{t-1}}(\psi)$. For $i=1,...,N$, set
\begin{align*}
m^{(i)} & = a\psi^{(i)}+(1-a)\hat{\psi} \\
\hat{x}_{t}^{(i)} & = E(x_{t}|x_{t-1}=x_{t-1}^{(i)},\psi=\psi^{(k)})
\end{align*}
\item For $k=1,...,N$:
\begin{itemize}
\item Draw $I_{k}$, with $P(I_{k}=i) \propto w_{t-1}^{(i)}f_{N}(y_{t}|0,e^{\frac{\hat{x}_{t}^{(i)}}{2}}) $
\item Draw $\psi^{(k)}$ from $N(m^{(I_{k})},h^{2}\Sigma)$
\item Draw $x_{t}^{(k)}$ from $N(\alpha^{(k)}+\beta^{(k)}x_{t-1}^{(I_{k})},\tau^{2})$
\item Set  $\tilde{w}_{t}^{(k)} = \frac{f_{N}(y_{t}|0,e^{\frac{x_{t}^{(k)}}{2}})}{f_{N}(y_{t}|0,e^{\frac{\hat{x}_{t}^{(I_{k})}}{2}})}$
\end{itemize}
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $\hat{\pi}_{t}=p(x_{t}|y_{1:t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{(x_{t}^{(i)},\psi^{(i)})}$
\end{enumerate}
\end{itemize}
\end{algorithm}

@@

    x<-rnorm(N,alpha+beta*x,tau)
    w1<-w*dnorm(data[t],0,exp(x/2))

@@

    weight = w*dnorm(data[t],0,exp(x/2))
    k   = sample(1:N,size=N,replace=TRUE,prob=weight)
    x1   = rnorm(N,alpha+beta*x[k],tau)
    lw  = dnorm(data[t],0,exp(x1/2),log=TRUE)-dnorm(data[t],0,exp(x[k]/2),log=TRUE)
    w   = exp(lw)
    w   = w/sum(w)
    ESS  = 1/sum(w^2)

@@

  for(t in 1:length(data)){
    x_prev=x
    x_hat = alpha+beta*x #means of the transition from the previous particle 
    I_k = sample(1:N,prob=w*dnorm(data[t],0,sqrt(exp((x_hat)/2))))
    for(k in 1:N){
      x[k]=rnorm(size=1,mean=alpha+beta*x_prev[I_k],sd=tau)
      w[k]=dnorm(data[t],0,sqrt(exp((x[k])/2)))/dnorm(data[t],0,sqrt(exp((x_hat)/2)))
    }
    w = w/sum(w)
    ESS  = 1/sum(w^2)

@@

    vpar     = var(pars)
    ms          = a*pars+(1-a)*matrix(mpar,N,3,byrow=T)
    mus         = pars[,1]+pars[,2]*xs 
    weight      = w*dnorm(data[t],0,exp(mus/2))
    k           = sample(1:N,size=N,replace=T,prob=weight)
    ms1         = ms[k,] + matrix(rnorm(3*N),N,3)%*%chol(h2*vpar)
    xt          = rnorm(N,ms1[,1]+ms1[,2]*xs[k],exp(ms1[,3]/2))
    w           = dnorm(data[t],0,exp(xt/2))/dnorm(data[t],0,exp(mus[k]/2))
    w           = w/sum(w)
    ESS         = 1/sum(w^2)

>>>>

\begin{align}
y_{t}|x_{t} & \sim N(0,e^{x_{t}}) \\
x_{t}|x_{t-1} & \sim N(\alpha+\beta x_{t-1},\tau^2)
\end{align}

@@

In the Bootstrap Particle filter we basically use the model transition kernel and 
do resampling any time the effective sample size is smaller than a predetermined threshold.
\begin{algorithm} BPF for Stochastic Volatility Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarize $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ form $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item Draw $x_{t}^{(i)} \sim N(\alpha+\beta x_{t-1},\tau^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t-1})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{1:t-1})$
\item Set $w_{t}^{(i)} = w_{t-1}^{(i)}f_{N}(y_{t};0,e^{x_{t}}) \ \ i=1,...,N$ such that $\{(x_{t},w_{t})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{1:t})$
\item Normalize the weights: $w_{t}^{(i)}=\frac{\tilde{w}_{t}^{(i)}}{\sum_{j=1}^{N}(\tilde{w}_{t}^{(j)})}$
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $p(x_{t}|y_{1:t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{enumerate}
\end{itemize}
\end{algorithm}

If we instead want to sample from the optimal transition kernel, things get more complicated. 
It can be shown that the optimal transition kernel is
\[ m_{t}^{opt}(x_{t}|x_{t-1}) \propto f_{t}(y_{t}|x_{t})p(x_{t}|x_{t-1}) \propto exp\bigg[-\frac{1}{2\sigma^{2}}(x_{t}-\alpha-\beta x_{t-1})^{2}-\frac{e^{-x_{t}}}{2}y_{t}^{2}-\frac{x_{t}}{2}\bigg]\]
and it is not straightforward to sample from it. Therefore, we follow a common praxis consisting in approximating $m_{t}^{opt}$ by a Gaussian distribution obtained by linearising $e^{-x_{t}}$, i.e. $e^{-x_{t}}=e^{-x_{0}}(1+x_{0}-x+...)$. Taking $x_{0}=\mu^{\star}(x_{t-1}) := \alpha + \beta x_{t-1}$, i.e the mean of the Gaussian transition density $p(x_{t}|x_{t-1})$. After some calculations we would end up with a Gaussian distribution $N(\xi(x_{t-1}),\sigma^{2})$ with 
\[ \xi(x_{t-1}) = \mu^{\star}(x_{t-1}) + \frac{\sigma^{2}}{4}(y_{t}^{2}e^{-\mu^{\star}(x_{t-1})}-2)\].
With this transition kernel we can implement the Guided Particle Filter in the following way

\begin{algorithm} GPF for Stochastic Volatility Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarize $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ from $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item Draw $x_{t}^{(i)} \sim N(\alpha+\beta x_{t-1}+\frac{\tau^2}{4}\big[y_{t}^2e^{-(\alpha+\beta x_{t-1})}-2\big],\tau^{2}) \ \ i=1,...,N$ such that $\{(x_{t},w_{t-1})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{1:t-1})$
\item Set $w_{t}^{(i)} = w_{t-1}^{(i)}\frac{f_{N}(y_{t};0,e^{x_{t}})f_{N}(x_{t};\alpha+\beta x_{t-1},\tau^{2})}{f_{N}(x_{t};\alpha+\beta x_{t-1}+\frac{\tau^2}{4}\big[y_{t}^2e^{-(\alpha+\beta x_{t-1})}-2\big],\tau^{2})} \ \ i=1,...,N$ such that $\{(x_{t},w_{t})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{1:t})$
\item Normalize the weights: $w_{t}^{(i)}=\frac{\tilde{w}_{t}^{(i)}}{\sum_{j=1}^{N}(\tilde{w}_{t}^{(j)})}$
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $p(x_{t}|y_{1:t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{enumerate}
\end{itemize}
\end{algorithm}

The Auxiliary Particle Filter with optimal transition kernel uses the same transition kernel of the Guided Particle Filter and we decided to use an auxiliary function $g(x_{t-1})=E(x_{t}|x_{t-1})=\alpha+\beta x_{t-1}$.

\begin{algorithm}APF with optimal transition kernel for Stochastic Volatility Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ from $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw $I_{k}$ with $P(I_{k}=i) \propto w_{t-1}^{(i)}f_{N}(y_{t}|0,e^{\alpha+\beta x_{t-1}}) $
\item $x_{t}^{(i)} \sim N(\alpha+\beta x_{t-1}^{(I_{k})}+\frac{\tau^2}{4}\big[y_{t}^2e^{-(\alpha+\beta x_{t-1}^{(I_{k})})}-2\big],\tau^{2})$ for $i=1,...N$
\item Set  $\tilde{w}^{(i)}_{t} = N^{-1}$ for $i=1,...,N$
\end{enumerate}
\item otherwise
\begin{enumerate}
\item Draw $x_{t}^{(i)} \sim N(\alpha+\beta x_{t-1}+\frac{\tau^2}{4}\big[y_{t}^2e^{-(\alpha+\beta x_{t-1})}-2\big],\tau^{2}) \ \ i=1,...,N$ for $i=1,...N$
\item Set $\tilde{w}^{(i)}_{t}=w^{(i)}_{t-1}f_{N}(y_{t};0,\alpha+\beta x_{t})$ for $i=1,...,N$
\end{enumerate}
\item Normalize the weights: $w_{t}^{(i)}=\frac{\tilde{w}_{t}^{(i)}}{\sum_{j=1}^{N}(\tilde{w}_{t}^{(j)})}$
\end{enumerate}
\end{itemize}
\end{algorithm}

Finally we 
\begin{algorithm} LWF for Random Walk plus Noise Model
\begin{itemize}
\item Initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ from $N(m_{0},C_{0})$, $(\alpha^{(1)},...,\alpha^{(N)})$ from $N(e_{\alpha},v_\alpha)$ and
$(\beta^{(1)},...,\beta^{(N)})$ from $N(e_{\beta},v_\beta)$ $({\tau^{2}}^{(1)},...,{\tau^{2}}^{(N)})$ from $IG(\nu_{0}/2,\nu_{0}\lambda/2)$. Set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$ and chose the appropriate value for $\delta$. Therefore $\psi^{(i)}=({\alpha}^{(i)},{\beta}^{(i)},{\tau^{2}}^{(i)})$, and
$\hat{\pi}_{0}=p(x_{0}|y_{0})=\sum_{i=1}^{N}w_{0}^{(i)}\delta_{(x_{0}^{(i)},\psi^{(i)})}$
\item For $t=1,...,n$:.
\begin{enumerate}
\item Compute $\overline{\psi}=E_{\hat{\pi}_{t-1}}(\psi)$ and $\Sigma=V_{\hat{\pi}_{t-1}}(\psi)$. For $i=1,...,N$, set
\begin{align*}
m^{(i)} & = a\psi^{(i)}+(1-a)\hat{\psi} \\
\hat{x}_{t}^{(i)} & = E(x_{t}|x_{t-1}=x_{t-1}^{(i)},\psi=\psi^{(k)})
\end{align*}
\item For $k=1,...,N$:
\begin{itemize}
\item Draw $I_{k}$, with $P(I_{k}=i) \propto w_{t-1}^{(i)}f_{N}(y_{t}|0,e^{\frac{\hat{x}_{t}^{(i)}}{2}}) $
\item Draw $\psi^{(k)}$ from $N(m^{(I_{k})},h^{2}\Sigma)$
\item Draw $x_{t}^{(k)}$ from $N(\alpha^{(k)}+\beta^{(k)}x_{t-1}^{(I_{k})},\tau^{2})$
\item Set  $\tilde{w}_{t}^{(k)} = \frac{f_{N}(y_{t}|0,e^{\frac{x_{t}^{(k)}}{2}})}{f_{N}(y_{t}|0,e^{\frac{\hat{x}_{t}^{(I_{k})}}{2}})}$
\end{itemize}
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $\hat{\pi}_{t}=p(x_{t}|y_{1:t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{(x_{t}^{(i)},\psi^{(i)})}$
\end{enumerate}
\end{itemize}
\end{algorithm}

@@

    x<-rnorm(N,alpha+beta*x,tau)
    w1<-w*dnorm(data[t],0,exp(x))

@@

    weight = w*dnorm(data[t],0,exp(x))
    k   = sample(1:N,size=N,replace=TRUE,prob=weight)
    x1   = rnorm(N,alpha+beta*x[k],tau)
    lw  = dnorm(data[t],0,exp(x1),log=TRUE)-dnorm(data[t],0,exp(x[k]),log=TRUE)
    w   = exp(lw)
    w   = w/sum(w)
    ESS  = 1/sum(w^2)

@@

  for(t in 1:length(data)){
    x_prev=x
    x_hat = alpha+beta*x #means of the transition from the previous particle 
    I_k = sample(1:N,prob=w*dnorm(data[t],0,sqrt(exp((x_hat)/2))))
    for(k in 1:N){
      x[k]=rnorm(size=1,mean=alpha+beta*x_prev[I_k],sd=tau)
      w[k]=dnorm(data[t],0,sqrt(exp((x[k]))))/dnorm(data[t],0,sqrt(exp((x_hat))))
    }
    w = w/sum(w)
    ESS  = 1/sum(w^2)

@@

    vpar     = var(pars)
    ms          = a*pars+(1-a)*matrix(mpar,N,3,byrow=T)
    mus         = pars[,1]+pars[,2]*xs 
    weight      = w*dnorm(data[t],0,exp(mus))
    k           = sample(1:N,size=N,replace=T,prob=weight)
    ms1         = ms[k,] + matrix(rnorm(3*N),N,3)%*%chol(h2*vpar)
    xt          = rnorm(N,ms1[,1]+ms1[,2]*xs[k],exp(ms1[,3]/2))
    w           = dnorm(data[t],0,exp(xt))/dnorm(data[t],0,exp(mus[k]))
    w           = w/sum(w)
    ESS         = 1/sum(w^2)
